---
title: "Lab 6"
output: html_notebook
author: Kashmira De Saram
---
### packages needed
```{r}

# packages
library(tidyverse)
library(tidycensus)
library(sf)
library(ggplot2)


options(tigris_class = "sf")
options(tigris_use_cache = TRUE)

setwd("C:/Users/folder/ex")
```
### Data
- Get from (https://www.socialexplorer.com/data/metadata/)
- I used the Single Female and Male households census data with ACS 2019 (5-year estimates)
```{r}
Mgomery_household_2019 <- get_acs(geography = "tract", 
     variables = c("total_hhtype" = "B11012_001", # All household types
                   "Female_hh" = "B11012_008", # All Female households, no spouse
                   "Male_hh" = "B11012_013", # All Male households, no spouse 
                   "med_hh_inc" = "B19013_001" # Median household income
                   ), 
     year = 2019,
     survey = "acs5",
     state = c(24), 
     county = c(031), 
     geometry = TRUE,
     output = "wide") 

```
This is for 2014
```{r}
Mgomery_household_2014 <- get_acs(geography = "tract", 
     variables = c("total_hhtype" = "B11012_001", # All household types
                   "Female_hh" = "B11012_010", # All Female households, no spouse
                   "Male_hh" = "B11012_007", # All Male households, no spouse 
                   "med_hh_inc" = "B19013_001" # Median household income
                   ),      
     year = 2014,
     survey = "acs5",
     state = c(24), 
     county = c(031),
     geometry = FALSE, 
     output = "wide") # need this
```
### Next convert to cvs, this will show all the computed data

```{r}

st_write(Mgomery_household_2014, "Mgomery_household_2014.csv") 
# geometry is false!
```
### Finding the percentage of single female and male households compared to all other household types
```{r}
# Compute percentage of Female and Male (with np spouse) households in 2019
Mgomery_household_2019$Both_households <- (Mgomery_household_2019$Female_hhE + Mgomery_household_2019$Male_hhE) / Mgomery_household_2019$total_hhtypeE

Mgomery_household_2019$Other_households <- 1- Mgomery_household_2019$Both_household


```
```{r}
# Compute percentage of Female and Male (with np spouse) households in 2014
Mgomery_household_2014$Both_households <- (Mgomery_household_2014$Female_hhE + Mgomery_household_2014$Male_hhE)/ Mgomery_household_2014$total_hhtypeE

Mgomery_household_2014$Other_households <- 1- Mgomery_household_2014$Both_households
```

### Merging the two tables with a right_join
```{r mergeanddiff}
## Merge the two time periods
Mgomery_household <- right_join(Mgomery_household_2019, Mgomery_household_2014, 
                              by="GEOID",
                              suffix=c(".19",".14"))
```

### computing the differece in single female and male households (individually) between the 5 year estimates
### Also, reproject to 3857 (web mercator)
```{r}

# Compute differences in households
Mgomery_household$Female_hhE <- Mgomery_household$Female_hhE.19 -
        Mgomery_household$Female_hhE.14

Mgomery_household$Male_hhE <- Mgomery_household$Male_hhE.19 -
        Mgomery_household$Male_hhE.14

# Compute difference in median household income
Mgomery_household$mhhi_diff <- Mgomery_household$med_hh_incE.19 - Mgomery_household$med_hh_incE.14


Mgomery_household <- st_transform(Mgomery_household, 3857) # reproject into web-mercator  
```
### write the cvs to geojson
```{r}

st_write(Mgomery_household_2019, "Mgomery_household_diff.geojson")
```

```


